Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-01-12 response-to-reviews.Rmd

**We agree that more caution is warrented on comparing canopy vs understory responses to warming, particularly because understory responses are highly uncertain. We have reworded relevant text as follows:**

- **"Implications: Global Change Responses : Warming" section, 4th par (previous lines 793-797: "While it is currently difficult to predict whether canopy or understory photosynthesis is likely to be more severely affected by higher $T_{air}$ (see section 4.2), *limited tree-ring evidence* indicates that understory trees *can* exhibit greater reductions in growth during unusually hot growing seasons [section 5.1, Fig 4b, @rollinsonClimateSensitivityUnderstory2020]."**

- **"Implications: Global Change Responses : Warming" section, final par (previous lines 803-806: "We expect that the tallest trees will be increasingly prone to hydraulic failure and damaging or lethal $T_{leaf}$'s, *while much larger uncertainty remains surrounding the resoponses of understory trees*."**

- **Conclusions section, first par (text specifically cited in the comment above): "The preponderance of available data suggest that large canopy trees are the most vulnerable to warming when water is limited. *While limited evidence suggests that understory trees may be more vulnerable to chronic warming stress under relatively mesic conditions*, more research is needed.**

### 2022-01-12 vertical-thermal-review.Rmd
edit conclusions - issue #90

Across vertical gradients, directional trends in the biophysical environment and leaf traits are the rule, driving variation in the physiology and ecology that have these as their underpinnings (Fig. `r fig_schematic`).
However, there remains a lot of uncertainty as to how the temperature sensitivities of metabolism and woody growth vary across these vertical gradients.
The preponderance of available data suggest that while large canopy trees are the most vulnerable to warming when water is limited, understory trees may be more vulnerable under more mesic conditions, but more research is needed.

>>>>

Across vertical gradients, directional trends in the biophysical environment and leaf traits are the rule, driving variation in the physiology and ecology that have these as their underpinnings (Fig. `r fig_schematic`).
However, there remains a lot of uncertainty as to how the temperature sensitivities of metabolism and woody growth vary across these vertical gradients.
The preponderance of available data suggest that large canopy trees are the most vulnerable to warming when water is limited.
While limited evidence suggests that understory trees may be more vulnerable to chronic warming stress under relatively mesic conditions, more research is needed.
